---
title: "Take-home Exercise 1: Application of Spatial Point Patterns Analysis to discover the geographical distribution of functional and non-function water points in Osun State, Nigeria"
date: "30 January 2023"
date-modified: "`r Sys.Date()`"
execute:
  echo: true
  eval: true
  warning: false
editor: visual
---

# Setting the Scene

Water is an important resource to mankind. Clean and accessible water is critical to human health. It provides a healthy environment, a sustainable economy, reduces poverty and ensures peace and security. Yet over 40% of the global population does not have access to sufficient clean water. By 2025, 1.8 billion people will be living in countries or regions with absolute water scarcity, according to UN-Water. The lack of water poses a major threat to several sectors, including food security. Agriculture uses about 70% of the world\'s accessible freshwater.

Developing countries are most affected by water shortages and poor water quality. Up to 80% of illnesses in the developing world are linked to inadequate water and sanitation. Despite technological advancement, providing clean water to the rural community is still a major development issues in many countries globally, especially countries in the Africa continent.

To address the issue of providing clean and sustainable water supply to the rural community, a global [Water Point Data Exchange (WPdx)](https://www.waterpointdata.org/about/) project has been initiated. The main aim of this initiative is to collect water point related data from rural areas at the water point or small water scheme level and share the data via WPdx Data Repository, a cloud-based data library. What is so special of this project is that data are collected based on [WPDx Data Standard](https://www.waterpointdata.org/wp-content/uploads/2021/04/WPDx_Data_Standard.pdf).

# Objectives

Geospatial analytics hold tremendous potential to address complex problems facing society. In this study, we are tasked to apply appropriate spatial point patterns analysis methods to discover the geographical distribution of functional and non-function water points and their co-locations if any in Osun State, Nigeria.

# The Task

The specific tasks of this take-home exercise are as follows:

## Exploratory Spatial Data Analysis (ESDA)

-   Derive kernel density maps of functional and non-functional water points. Using appropriate tmap functions,

-   Display the kernel density maps on openstreetmap of Osub State, Nigeria.

-   Describe the spatial patterns revealed by the kernel density maps. Highlight the advantage of kernel density map over point map.

## **Second-order Spatial Point Patterns Analysis**

With reference to the spatial point patterns observed in ESDA:

-   Formulate the null hypothesis and alternative hypothesis and select the confidence level.

-   Perform the test by using appropriate Second order spatial point patterns analysis technique.

-   With reference to the analysis results, draw statistical conclusions.

## Spatial Correlation Analysis

In this section, we are required to confirm statistically if the spatial distribution of functional and non-functional water points are independent from each other.

-   Formulate the null hypothesis and alternative hypothesis and select the confidence level.

-   Perform the test by using appropriate Second order spatial point patterns analysis technique.

-   With reference to the analysis results, draw statistical conclusions.

# The Data

## **Aspatial data**

For the purpose of this assignment, data from [WPdx Global Data Repositories](https://www.waterpointdata.org/access-data/) will be used. There are two versions of the data. They are: WPdx-Basic and WPdx+. This take home assignment will use the **WPdx+** data set.

## **Geospatial data**

This study will focus of Osun State, Nigeria. The state boundary GIS data of Nigeria can be downloaded either from The [Humanitarian Data Exchange](https://data.humdata.org/) portal or [geoBoundaries](https://www.geoboundaries.org/).

# Install and Load Packages

```{r}

```

# Data Import

```{r}

```

# 
